A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data

### 2023-08-14 tidyindex.qmd
outline insights from the gggi plot

When economy and education are given a greater weight, the variation in index values and ranking highlights countries that performs relatively better (moving right) or worse (moving left) in these two single dimensions. For example, the index moves to the left for Ethiopia and Senegal when economy is given a higher weight. This reveals the economy as a weakness for these two countries compared to their similarly-ranked peers.

When health receives a higher weight, the impact on the index value is minimal.

Notably, increasing the weight in politics leads to a pronounced decrease of index value for almost all the countries. Given the index value has a direct interpretation on how much of the gender gap has been closed, such a weight variation can cast a negative light on the progression to gender parity. Analysts need to further investigate the construction of the politics dimension and consider whether this is desirable for the index.

@@

* Bangladesh uses politics to compensate economy
* increase economy -> 
* increase education -> 
* increase health -> shrink the range of index (bring top and bottom country closer)
* increase politics -> all the country has index value smaller

@@

>>>>

* Bangladesh 
* when decreasing the weight on politics, difference between countries narrower. 
* increase politics -> all the country has lower index value

<!-- Notably, increasing the weight in politics leads to a pronounced decrease of index value for almost all the countries. Given the index value has a direct interpretation on how much of the gender gap has been closed, such a weight variation can cast a negative light on the progression to gender parity. Analysts need to further investigate the construction of the politics dimension and consider whether this is desirable for the index. -->

@@

@@

<!-- When economy and education are given a greater weight, the variation in index values and ranking highlights countries that performs relatively better (moving right) or worse (moving left) in these two single dimensions. For example, the index moves to the left for Ethiopia and Senegal when economy is given a higher weight. This reveals the economy as a weakness for these two countries compared to their similarly-ranked peers.  -->

### 2023-08-14 tidyindex.tex
outline insights from the gggi plot

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne,
Victoria, Australia\\ \textsuperscript{2} Institute of
Statistics, University of Natural Resources and Life
Sciences, Vienna, Austria\\ \textsuperscript{3} Department of
Mathematical Sciences, BNU-HKBU United International College, Zhuhai,
Guangdong, China
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep 
software design
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners, breakable, interior hidden]}{\end{tcolorbox}}\fi

@@

When economy and education are given a greater weight, the variation in
index values and ranking highlights countries that performs relatively
better (moving right) or worse (moving left) in these two single
dimensions. For example, the index moves to the left for Ethiopia and
Senegal when economy is given a higher weight. This reveals the economy
as a weakness for these two countries compared to their similarly-ranked
peers.

When health receives a higher weight, the impact on the index value is
minimal.

Notably, increasing the weight in politics leads to a pronounced
decrease of index value for almost all the countries. Given the index
value has a direct interpretation on how much of the gender gap has been
closed, such a weight variation can cast a negative light on the
progression to gender parity. Analysts need to further investigate the
construction of the politics dimension and consider whether this is
desirable for the index.

\begin{itemize}
\tightlist
\item
  Bangladesh uses politics to compensate economy
\item
  increase economy -\textgreater{}
\item
  increase education -\textgreater{}
\item
  increase health -\textgreater{} shrink the range of index (bring top
  and bottom country closer)
\item
  increase politics -\textgreater{} all the country has index value
  smaller
\end{itemize}

>>>>

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne,
Victoria, Australia\\ \textsuperscript{2} Institute of
Statistics, University of Natural Resources and Life
Sciences, Vienna, Austria\\ \textsuperscript{3} Department of
Mathematical Sciences, BNU-HKBU United International College, Zhuhai,
Guangdong, China
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep 
software design
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, frame hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi

@@

\begin{itemize}
\tightlist
\item
  Bangladesh
\item
  when decreasing the weight on politics, difference between countries
  narrower.
\item
  increase politics -\textgreater{} all the country has lower index
  value
\end{itemize}
